Compositional Certified Resource Bounds

### 2015-04-13 paper/main.tex
add note on LP certificates, move figures

We present the rules for a subset of the
expressions of Clight.  Clight is the first intermediate
language of the CompCert compiler~\cite{CompcertCACM}, it
is a cleaner C with a unified looping construct and
side-effect free expressions.  We reuse most of CompCert's
syntax but instrument the semantics with a resource metric
$M$ that accounts for the cost (an arbitrary rational number)
of each step in the operational semantics.  For example, $M_e(\code{exp})$ is the cost of
evaluating the expression \code{exp}; the rationals $M_f$
and $M_r$ account respectively for the cost of a call to the
function $f$ and the cost of returning from it.
More details on this cost-aware semantics are provided in \pref{sec:soundness}.

@@

A notable advantage of the LP-based approach compared to
SMT-solver--based techniques is that a satisfying assignment is
a proof certificate instead of a counter example.

@@

More precisely, the technique consists of the following steps. First,
a program $P$ that fails to be analyzed automatically is enriched by
auxiliary variables $\vec x$ and assertions to form a program
$P_l(\vec x)$.  Second, an initial value $\vec X(\sigma)$ for the
logical variables is selected so that:
$$
\forall n\, \sigma\, \sigma'.\,
  (\sigma, P_l(\vec X(\sigma))) \Downarrow_n \sigma'
    {\implies}
  \exists n' {\le} n.\,
    (\sigma, P) \Downarrow_{n'} \sigma'.
$$
Since the annotated program and the original one are usually
syntactically close, the proof of this result goes
by simple induction on the resource-aware evaluation judgement.
Third, using existing automation tools, a bound $B(\vec x)$ for
$P_l(\vec x)$ is derived.  Finally this bound, instantiated with
$\vec X$, gives the final resource bound for the program $P$.

>>>>

We present the rules for a subset of the
expressions of Clight.  Clight is the first intermediate
language of the CompCert compiler~\cite{CompcertCACM}, it
is a cleaner C with a unified looping construct and
side-effect free expressions.  We reuse most of CompCert's
syntax but instrument the semantics with a resource metric
$M$ that accounts for the cost (an arbitrary rational number)
of each step in the operational semantics.  For example, $M_e(\code{exp})$ is the cost of
evaluating the expression \code{exp}; the rationals $M_f$
and $M_r$ account respectively for the cost of a call to the
function $f$ and the cost of returning from it.
More details are provided in \pref{sec:soundness}.

@@

A notable advantage of the LP-based approach compared to
SMT-solver--based techniques is that a satisfying assignment is
a proof certificate instead of a counter example.  To provide
high-assurance bounds, this certificate can be checked
in linear time by a simple validator.

@@

More precisely, the technique consists of the following steps. First,
a program $P$ that fails to be analyzed automatically is enriched by
auxiliary variables $\vec x$ and assertions to form a program
$P_l(\vec x)$.  Second, an initial value $\vec X(\sigma)$ for the
logical variables is selected to satisfy the proposition:
$$
\forall n\, \sigma\, \sigma'.\,
  (\sigma, P_l(\vec X(\sigma))) \Downarrow_n \sigma'
    {\implies}
  \exists n' {\le} n.\,
    (\sigma, P) \Downarrow_{n'} \sigma'.
$$
Since the annotated program and the original one are usually
syntactically close, the proof of this result goes
by simple induction on the resource-aware evaluation judgement.
Third, using existing automation tools, a bound $B(\vec x)$ for
$P_l(\vec x)$ is derived.  Finally this bound, instantiated with
$\vec X$, gives the final resource bound for the program $P$.
